Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-07-28 Moltres.tex
Complete MSRE comparison.

\title{Introduction to Moltres: an Application for Simulation of Molten Salt Reactors}
\author{Alexander Lindsay, Kathryn Huff}

@@

\Cref{fig:flux_comparison} compares the fast and thermal neutron fluxes at the
reactor mid-plane ($z=H/2$). Local thermal flux growth and fast flux decay in
moderator regions and visa versa in fuel regions are even more apparent. The
larger fast flux magnitude is attributable to using \gls{BOL} fuel
composition. As the fuel salt reaches an equilibrium composition, the fast to
thermal flux ratio is expected to decrease.

\begin{figure}[htpb]
  \centering
  \includegraphics{radial_group_comparison.eps}
  \caption{The thermal and fast flux profiles are compared at the core mid-plane
    ($z=H/2$) in this 2-D cylindrical axisymmetric model ($r=0$ is radial center
    of core). }
  \label{fig:flux_comparison}
\end{figure}

@@

\Cref{fig:temp_compare} shows a comparison between Moltres predicted temperature
profiles with cosinusoidal gamma heating and \gls{MSRE} design models
\cite{briggs_molten-salt_1964} along the reactor center-line (r = 0).

\begin{figure}[htpb]
    \centering
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{fuel_vs_moderator_axial_sweep.eps}
      \caption{Moltres predicted axial temperature profiles near r = 0}
      \label{fig:moltres_axial_temps}
    \end{subfigure}
    %\quad
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{msre_axial_temps.png}
      \caption{\gls{MSRE} design axial temperature profiles near r = 0
        \cite{briggs_molten-salt_1964}}
      \label{fig:msre_axial_temps}
    \end{subfigure}
    \caption{Comparison of \gls{MSRE} design and Moltres predicted axial
      temperature profiles}
    \label{fig:temp_compare}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{msre_radial_fluxes.png}
  \caption{\gls{MSRE} design radial flux profiles at reactor mid-plane
    \cite{briggs_molten-salt_1964}}
  \label{fig:msre_radial_fluxes}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{msre_axial_fluxes.png}
  \caption{\gls{MSRE} design axial flux profiles at r = 0
    \cite{briggs_molten-salt_1964}}
  \label{fig:msre_axial_fluxes}
\end{figure}

@@

This work introduces the open source \gls{MSR} simulation code Moltres. Moltres
solves arbitrary-group neutron diffusion, temperature, and precursor governing
equations in anywhere from one to three dimensions and can be deployed on an
arbitrary number of processing units. The model problem presented here has a
2D-axisymmetric geometry with heterogeneous group constants for fuel and
moderator regions. Fuel volume fraction and fuel salt composition are based on
the \gls{MSRE}. Neutron fluxes show expected cosinusoidal shapes in radial and
axial directions with visible striations between fuel and moderator regions. The
fast group flux is enhanced in fuel regions while the thermal group flux in
enhanced in moderator regions. Due to advection the temperature profile
increases monotonically in the direction of salt flow. The role of advection is
also seen in precursor concentrations. Long lived precursors exhibit maximum
concentrations at the core outlet. As the decay constant increases across
precursor groups the maximum concentrations moves towards the reactor center
where the precursor production rate is maximum. Future Moltres work includes
generating a high-fidelity 3D model as well as investigating various transient
accident scenarios.

>>>>

\title{Introduction to Moltres: an Application for Simulation of Molten Salt Reactors}
\author{Alexander Lindsay, Gavin Ridley, Kathryn Huff}

@@

@@

\Cref{fig:temp_compare} shows a comparison between Moltres predicted temperature
profiles with cosinusoidal gamma heating and \gls{MSRE} design models
\cite{briggs_molten-salt_1964} in the hottest channel and adjacent graphite. The
profiles are in good agreement. In the \gls{MSRE} design, the moderator
temperature at the reactor inlet is about 20 degrees Faerenheit larger than the
fuel temperature, whereas the temperatures are about the same in the Moltres
model. This difference can perhaps be attributed to the neglection of axial
conduction in the \gls{MSRE} design model \cite[p. 99]{briggs_molten-salt_1964}.

\begin{figure}[htpb]
    \centering
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{fuel_vs_moderator_axial_sweep.eps}
      \caption{Moltres predicted axial temperature profiles in hottest channel
        and adjacent graphite}
      \label{fig:moltres_axial_temps}
    \end{subfigure}
    %\quad
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{msre_axial_temps.png}
      \caption{\gls{MSRE} design axial temperature profiles in hottest channel
        and adjacent graphite stringer \cite[p. 99]{briggs_molten-salt_1964}}
      \label{fig:msre_axial_temps}
    \end{subfigure}
    \caption{Comparison of \gls{MSRE} design and Moltres predicted axial
      temperature profiles}
    \label{fig:temp_compare}
\end{figure}

\Cref{fig:moltres_radial_fluxes} compares the fast and thermal neutron fluxes at the
reactor mid-plane ($z=H/2$). Local thermal flux growth and fast flux decay in
moderator regions and visa versa in fuel regions are apparent. The flux
magnitudes are in good agreement with the magnitudes from the \gls{MSRE} design
calculations \cite[p. 92]{briggs_molten-salt_1964} as shown in
\cref{fig:msre_radial_fluxes}. The peak fast to thermal flux ratio is
approximately 3.5 in the \gls{MSRE} design calculation as opposed to a ratio of
3 for the Moltres calculation. The depression in the thermal flux in
\cref{fig:msre_radial_fluxes} is caused by control rod thimbles and an extra
volume of surrounding fuel that is not included in the Moltres calculations.

\begin{figure}[htpb]
    \centering
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{moltres_radial_fluxes.eps}
      \caption{The thermal and fast flux profiles at the core mid-plane
        ($z=H/2$) for Moltres 2-D cylindrical axisymmetric model ($r=0$ is radial center
        of core)}
      \label{fig:moltres_radial_fluxes}
    \end{subfigure}
    %\quad
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{msre_radial_fluxes.png}
      \caption{The thermal and fast flux profiles at the core mid-plane
        ($z=H/2$) for \gls{MSRE} design model \cite{briggs_molten-salt_1964}
        ($r=0$ is radial center of core)}
      \label{fig:msre_radial_fluxes}
    \end{subfigure}
    \caption{Comparison of \gls{MSRE} design and Moltres predicted radial
      flux profiles}
    \label{fig:radial_fluxes_compare}
\end{figure}

\Cref{fig:axial_fluxes_compare} compares the axial flux profiles calculated by
Moltres and the \gls{ORNL} \gls{MSRE} design model. Note that the radius for the
plots are chosen to correspond to the peak of the thermal flux in both cases;
for the \gls{ORNL} calculations this is 8.4 inches from the core center-line
because of the effect of the control rod thimbles and extra fuel along the
center-line. Once again, the plots are in decent agreement. The \gls{ORNL}
calculations include the lower and upper plena which are not included in this
report's Moltres model. Consequently, the plot in \cref{fig:msre_axial_fluxes}
extends to lower and higher z-values than
\cref{fig:moltres_axial_fluxes}. Additionally, the deviation of the thermal flux
from a sinusoidal shape in \cref{fig:msre_axial_fluxes} is due to absorption in
the plena.

\begin{figure}[htpb]
    \centering
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{moltres_axial_fluxes.eps}
      \caption{Moltres axial flux profiles along core center line}
      \label{fig:moltres_axial_fluxes}
    \end{subfigure}
    %\quad
    \begin{subfigure}[b]{.5\textwidth}
      \includegraphics[width=\maxwidth{\textwidth}]{msre_axial_fluxes.png}
      \caption{\gls{MSRE} design axial flux profiles 8.4 inches from the core
        center line \cite{briggs_molten-salt_1964}}
      \label{fig:msre_axial_fluxes}
    \end{subfigure}
    \caption{Comparison of \gls{MSRE} design and Moltres predicted axial
      flux profiles}
    \label{fig:axial_fluxes_compare}
\end{figure}

@@

This work introduces the open source \gls{MSR} simulation code Moltres. Moltres
solves arbitrary-group neutron diffusion, temperature, and precursor governing
equations in anywhere from one to three dimensions and can be deployed on an
arbitrary number of processing units. The 2D-axisymmetric and 3D models
presented here employ heterogeneous group constants for fuel and moderator
regions generated with SCALE. Fuel volume fraction and fuel salt composition are
based on the \gls{MSRE}. Neutron fluxes show expected cosinusoidal shapes in
radial and axial directions with visible striations between fuel and moderator
regions. The fast group flux is enhanced in fuel regions while the thermal group
flux in enhanced in moderator regions. Due to advection the temperature profile
in the fuel increases monotonically in the direction of salt flow, while the
moderator temperature exhibits a maximum between the mid-plane and outlet. The
role of advection is also seen in precursor concentrations. Long lived
precursors exhibit maximum concentrations at the core outlet. As the decay
constant increases across precursor groups the maximum concentrations moves
towards the reactor center where the precursor production rate is
maximum. Results from 2D-axisymmetric and 3D models show good qualitative
agreement. Moreover, Moltres results compare favorably with the actual design
calculations of the \gls{MSRE}. Strong parallel scaling was demonstrated on a
typical Moltres model problem. Future Moltres publications will highlight
transient simulation cases investigating control rod ejection, single channel
blockage, loss of flow, and loss of secondary cooling.
